Code to reproduce results in _Semi-parametric generalized linear model for

### 2021-08-29 ManuscriptCode/Manuscript.Rmd
Update simulation study

 * simulate data with either a beta-binomial or a q-power backbone distribution that follows a semi-parametric logistic regression model
 * fit the data using our semi-parametric regression model, beta-binomial regression (potentially misspecified distribution), and GEE (marginal only)

@@

The dose levels will have mean 0.2, 0.4, 0.6, and 0.8, and intra-cluster correlation 0.4.

```{r SimOne}
generate_data <- function(pdf.fun, maxClusterSize = 10, aveClusterSize = 9,
                          doseLevels = 4, nPerDose = 25){
  # generate binomially distributed frequencies for each cluster-size
  ss <- expand.grid(Trt = 1:doseLevels, ClusterSize = 0:maxClusterSize) %>%
    group_by(Trt) %>%
    mutate(Freq = rmultinom(1, size=nPerDose, 
                            prob=dbinom(ClusterSize, size=maxClusterSize, 
                                        prob=aveClusterSize / maxClusterSize))) %>%
    filter(Freq > 0)

  res <- ran.CBData(sample.sizes = ss, 
                p.gen.fun = function(g) 0.2*g,
                rho.gen.fun = function(g) 0.4,
                pdf.fun = pdf.fun)
  res$Z <- rnorm(nrow(res), mean=0, sd=1)  
  res
}
```

@@

```{r AnalyzeOne}
run_analysis <- function(dat){
  #GEE
  dat_long <- unwrap(dat)
  res_gee <- geeglm(Resp ~ Trt + Z, data=dat_long, id = ID, family=binomial,
                 corstr = "exch")
  #Beta-binomial
  res_bb <- vgam(cbind(NResp, ClusterSize - NResp) ~ Trt + Z, weights = Freq,
                 data=dat, family =  betabinomial(lmu="logitlink"))
  #SPGLM
  op <- options(warn=-1)
  res_spglm <- spglm(cbind(NResp, ClusterSize - NResp) ~ Trt + Z, data=dat,
                   weights  = Freq)
  options(op)

@@

```{r ActualSim}
R <- 50

# beta-binomial PDF
simBB.list <- 
  replicate(R,
            {simd <- generate_data(betabin.pdf);
             run_analysis(simd)}, 
            simplify = FALSE)

# q-power PDF
simQP.list <- 
  replicate(R,
            {simd <- generate_data(qpower.pdf);
             run_analysis(simd)}, 
            simplify = FALSE)

simBB <- bind_rows(simBB.list, .id="ID")
simQP <- bind_rows(simQP.list, .id="ID")

@@

```{r Compare}
comb <- bind_rows(`Beta-Binomial` = simBB, `q-Power`= simQP, .id="Generator")

# get correct values of parameters on a logit-model scale
mns <- c(0.2, 0.4, 0.6, 0.8)
logit.mns <- log(mns/(1-mns))
correct <- tibble(Term = c("(Intercept)","Trt2","Trt3", "Trt4","Z"),
                  Correct = c(logit.mns[1], logit.mns[-1] - logit.mns[1], 0))
comb <- left_join(comb, correct, by="Term")

>>>>

 * simulate data with four settings:

  1. Both the beta-binomial and semi-parameteric model are correct - a binomial backbone, which belongs to the exponential family
  2. Only the beta-binomial model is correct - generated from a beta-binomial distribution
  3. Only the semi-parametric model is correct - generated from the spglm, with a zero-inflated beta-binomial backbone
  4. Neither model is correct - generated from a zero-inflated beta-binomial distribution
 * fit the data using our semi-parametric regression model, beta-binomial regression (potentially misspecified distribution), and GEE (marginal only)

@@

We will have one binary predictor $X_1 \sim Bern(0.5)$ with $beta_1=0.5$ and one continuous predictor $X_2 \sim N(0,1)$ with $\beta_2 = 1$. The cluster size distribution will be binomial.

```{r SimOne}
beta_correct <- c(0, 0.5, 1)
sim_covariates <- function(n, maxClusterSize = 10, aveClusterSize = 7, beta=beta_correct){
  x1 <- rbinom(n, size=1, prob=0.5)
  x2 <- rnorm(n)
  lp <- cbind(1, x1, x2) %*% beta
  mns <- 1/(1+exp(-lp))

  cs <- rbinom(n, size=maxClusterSize, prob=aveClusterSize/maxClusterSize)

  data.frame(ClusterSize = cs, X1=x1, X2=x2, p_resp = mns)
}

# add random response counts using SPGLM model
add_response_spglm <- function(cov_data, q0){
  # cov_data should come from sim_covariates

  simres <- ran.spglm(n=cov_data$ClusterSize, means = cov_data$p_resp, q0 = q0)
  res <- cov_data
  res$NResp <- simres$NResp
  res$Trt <- "All"
  res$Freq <- 1
  class(res) <- c("CBData", class(res))
  res
}

# add random response counts from zero-inflated beta-binomial 
add_response_zoibb <- function(cov_data, pstr0, rho){
  # cov_data should come from sim_covariates

  res <- cov_data
  res$NResp <- rzoibetabinom(n = nrow(cov_data),
                             size = cov_data$ClusterSize,
                             prob = cov_data$p_resp,
                             rho = rho,
                             pstr0 = pstr0)
  res$Trt <- "All"
  res$Freq <- 1
  class(res) <- c("CBData", class(res))
  res
}
```

@@

```{r AnalyzeOne}
run_analysis <- function(dat){
  #GEE
  dat_long <- unwrap(dat)
  res_gee <- geeglm(Resp ~ X1 + X2, data=dat_long, id = ID, family=binomial,
                 corstr = "exch")
  #Beta-binomial
  res_bb <- vgam(cbind(NResp, ClusterSize - NResp) ~  X1 + X2, weights = Freq,
                 data=dat, family =  betabinomial(lmu="logitlink"))
  #SPGLM
  op <- options(warn=-1)
  res_spglm <- spglm(cbind(NResp, ClusterSize - NResp) ~ X1 + X2, data=dat,
                   weights  = Freq)
  options(op)

@@

```{r ActualSim1, warning=FALSE}
R <- 50

# Scenario 1: binomial PDF
sim1 <- 
  replicate(R,
            {simd <- sim_covariates(100) %>% add_response_zoibb(pstr=0, rho=0);
             run_analysis(simd)}, 
            simplify = FALSE)  %>%
   bind_rows(.id="ID")
```

```{r ActualSim2, warning=FALSE}
# Scenario 2: binomial PDF
sim2 <- 
  replicate(R,
            {simd <- sim_covariates(100) %>% add_response_zoibb(pstr=0, rho=0.3);
             run_analysis(simd)}, 
            simplify = FALSE) %>%
   bind_rows(.id="ID")

```{r ActualSim3, warning=FALSE}
# Scenario 3: SPGLM with zero-inflated betabinomial backbone
sim3 <- 
  replicate(R,
            {simd <- sim_covariates(100) %>% 
              add_response_spglm(q0 = dzoibetabinom(0:10, size=10, prob=0.4, rho = 0.3, pstr0 = 0.1));
             run_analysis(simd)}, 
            simplify = FALSE)  %>%
   bind_rows(.id="ID")

```{r ActualSim4, warning=FALSE}
# Scenario 4: zero-inflated betabinomial 
sim4 <- 
  replicate(R,
            {simd <- sim_covariates(100) %>% add_response_zoibb(pstr=0.1, rho=0.3);
             run_analysis(simd)}, 
            simplify = FALSE)  %>%
   bind_rows(.id="ID")

@@

```{r Compare}
comb <- bind_rows(`Both correct` = sim1, 
                  `BB correct`= sim2,
                  `SPGLM correct` = sim3, 
                  `Neither correct` = sim4,
                  .id="Generator")

# get correct values of parameters on a logit-model scale
correct <- tibble(Term = c("(Intercept)","X1","X2"),
                  Correct = beta_correct)
comb <- left_join(comb, correct, by="Term")
